### 2004-05-06 Documentation/part2.tex
latest version of part2

We tell \GLOBES\ that we want to use the built-in neutrino factory
flux, by setting the beam type variable to $1$. The target mass 
is $50\,\mathrm{kt}$ and a total of $ 4.264\cdot10^{21}$
muons is stored. The data is divided into 20 energy bins ranging from
$4\,\mathrm{GeV}$ to $50\,\mathrm{GeV}$. 
\begin{quote}
{\tt /* beam */}\\
{\tt \$beam\_type = 1}\\
{\tt \$target\_mass = 50}\\
{\tt \$time = 8}\\
{\tt \$stored\_muons = 5.33e+20}\\
{\tt \$numofbins = 20}\\
{\tt \$emin = 4.0}\\
{\tt \$emax = 50.0}
\end{quote}

Next a file has to be specified which contains the cross section we want to 
use.
\begin{quote}
{\tt /* cross section */}\\
{\tt cross(\#CC)<}\\
{\tt \hspace*{3ex}@cross\_file = XCC.dat}\\
{\tt >}
\end{quote}
The command {\tt cross} tells the parser that a cross section environment
begins and has the name {\tt \#CC}. These names are used to refer later on
to this specific environment and each name begins with a leading {\tt \#}.
Whenever we need to specify a cross section from here on we can use {\tt \#CC}
to refer to the cross section as given in the file {\tt XCC.dat}.

@@

The energy resolution function can easily be specified with
\begin{quote}
{\tt /* energy resolution */}\\
{\tt energy(\#MINOS)<}\\
{\tt \hspace*{3ex}@algorithm = 1}\\
{\tt \hspace*{3ex}@bins = 20}\\
{\tt \hspace*{3ex}@emin = 4.0}\\
{\tt \hspace*{3ex}@emax = 50.0}\\
{\tt \hspace*{3ex}@energy\_resolution = \{0.15,0.0,0.0\}}\\
{\tt >}
\end{quote}
The {\tt energy} command opens the corresponding energy environment and
the energy resolution function defined here is refered to by the name 
{\tt \#MINOS}. Out of several possibilities it employs algorithm $1$,
the simplest and fastest one. The actual energy resolution is specified
by the energy resolution variable which is a list of three elements. Each 
element is one parameter of a general resolution function as defined in 
equation ~\ref{eq:sigma_e}.

Now we define the one appearance and the corresponding disappearance 
channel of a neutrino factory: 
$\nu_e\rightarrow\nu_\mu$  and $\bar\nu_\mu\rightarrow\bar\nu_\mu$ 
($\mu^+$ stored).
\begin{quote}
{\tt /* channels */}\\
{\tt channel(\#appearance)<}\\
{\tt \hspace*{3ex}@channel = +: +: electron: muon: \#CC: \#MINOS}\\
{\tt >}\\
{\tt channel(\#disappearance)<}\\
{\tt \hspace*{3ex}@channel = +: -: muon: muon: \#CC: \#MINOS}\\
{\tt >}
\end{quote}
The first $\pm$ selects the polarity, \ie\ wether $\mu^+$ or $\mu^-$ are 
stored. The second $\pm$ determines whether neutrinos or anti-neutrinos are
taken from the flux table. The third position defines the initial flavour
and the forth position the final flavour, followed by the name of the cross
section and energy resolution function.

The last step is to assemble the channels into a rule.
\begin{quote}
{\tt /* rules */}\\
{\tt rule(\#rule1)<}\\
{\tt \hspace*{3ex}@signal = 0.45 @ \#appearance}\\
{\tt \hspace*{3ex}@signalerror = 0.001 : 0.0001}\\
{\tt \hspace*{3ex}@background = 1.0e-05 @ \#disappearance}\\
{\tt \hspace*{3ex}@backgroundcenter = 1 : 0.0}\\
{\tt \hspace*{3ex}@backgrounderror = 0.05 : 0.0001}\\
{\tt \hspace*{3ex}@errordim = 0}\\
{\tt \hspace*{3ex}@energy\_window = 4.0 : 50.0}\\
{\tt >}
\end{quote}
The signal variable says that the signal in our experiment is the 
channel named {\tt \#appearance} with an constant efficiency of $0.45$.
The signal error variable has two components: the first one is the
normalization error of the signal, here it is $0.1\%$. The second component
signifies the energy calibration error of the signal, whose definition 
can be found in section~\ref{sec:energy}. The background variable
specifies the composition of the beam background and this case is
given by the channel named {\tt \#disappearance}, \ie\ this are
the muons whose charge has been mis-identified. They enter the calculation
with an efficiency of $1\cdot 10^{-5}$. The background center variable allows
to rescale the total background contribution which is especially useful if
there are many background components. 
The background error variable is defined in the same way as the signal error
variable, \ie\ we have a $5\%$ background uncertainty.
The error dimension variable selects how
the systematical errors are treated according to table~\ref{tab:error_dim}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Source properties}
\label{sec:source}

\GLOBES\ handles only point sources, thus it is not possible
to study effects from the finite size of the neutrino production
region like in the Sun. Furthermore only one source for each experiment
is possible, this make it impossible\footnote{at least rather tricky} 
to simulate multiple source experiments like KamLAND. All neutrino
sources usable within \GLOBES\ therefore can be specified by the
flux spectra for each flavour and CP sign and the total luminosity
of the source. There are two principal ways to provide a flux:
use a built-in source or a file provided by the user. Built-in sources
as well as user defined fluxes are referred to internally by the
 {\tt \$beam\_type} variable. Certain values of  {\tt \$beam\_type} are
pre-defined and used for the built-in fluxes, \eg\ {\tt \$beam\_type = 1}
selects the neutrino factory flux.

The built-in source is muon decay, since the neutrino
spectra and flavour composition are known analytically in that case. For 
this sources the energy of the parent particle has to be specified
and the number of useful decays per year. Thus the sequence for setting
up a neutrino factory is
\begin{quote}
{\tt \$beam\_type = 1}\\
{\tt \$parent\_energy = 50.0}\\
{\tt \$stored\_muons = 5.33e+20}\\
\end{quote}
The values of {\tt \$parent\_energy} and {\tt emax} are always the same.

User defined fluxes are dealt with a little different at initialization
and the value of {\tt \$beam\_type} is automatically assigned. This is
necessary to avoid any problems with multiple experiments, since the fluxes
are read only once. Reading a flux file looks like
\begin{quote}
{\tt flux(\#dummy)<}\\
{\tt \hspace*{3ex}@flux\_file\_p = user\_file\_1.dat}\\
{\tt \hspace*{3ex}@flux\_file\_m = user\_file\_2.dat}\\
{\tt >}\\
{\tt \$target\_power = 4.0 }
\end{quote}

Since many sources have two modes of operation, it is possible to define
a pair of fluxes by using the polarity flag when loading the files. Usually
files connected in this way use the suffixes ``plus'' and ``minus'' in the 
file names. The individual members of the pair can be accessed by the polarity
variable. Also the built-in fluxes provide this choice.
The flux read by {\tt @flux\_file\_p} can be accesed by selecting positive
polarity, whereas the flux read by {\tt @flux\_file\_m} is selected by a 
negative value of the polarity.

@@

The format of the flux files is seven rows and
501 lines.
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
where the energy values are evenly spaced. 
For accessing fluxes at 
arbitrary energies linear interpolation is used. Once the energy leaves the
range of values given in the file $0.0$ is returned. The units used for the
fluxes are not specified and right now each flux has its own fudge factor
which is hard encoded in the file {\tt nufact.c}. The problem is that
in computing the normalization one has to know things like number of target
particles per unit mass asf. which is not always straightforward. Thus a fudge
factor will be useful in any case. Right now however this is not the only 
fudge factor since for some experiments the definition of their target mass
involves another fudge factor. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Oscillation properties}

@@

The most obvious property of the neutrino detection is the total
active mass of the target as already defined in section~\ref{sec:source} and
its value is defined by 
\begin{quote}
{\tt \$target\_mass = 50.0 }
\end{quote}
Furthermore the running time has to be specified by
\begin{quote}
{\tt \$time = 8.0 }
\end{quote}

@@

A cross section file is read by 
\begin{quote}
{\tt flux(\#name)<}\\
{\tt \hspace*{3ex}@cross\_file =user\_file\_1.dat}\\
{\tt >}
\end{quote}  
and can later be refered to by {\tt \#name}.

@@

Since it is rather cumbersome to numerically solve this double integral,
we split the two integrations. We first evaluate the integral over
$\hat{E}$, where the only terms containing $\hat{E}$ are
$k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
$ V_\beta(\hat{E}-E')$, and define:
\begin{eqnarray}
\label{eq:e_res} 
R_\beta^{\text{IT}}(E,E')\,\epsilon_\beta^{\text{IT}}(E')
 \equiv
\int d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
\,V_\beta(\hat{E}-E')\,, 
\end{eqnarray}

@@

@@

Based on the definition of channels it is possible to assemble the 
signal and background event numbers for each bin.  
The number of events in one channel and  in one bin\index{Bin} $i$ is given by
\begin{equation}
n_i^c=\int_{E_i-\Delta E/2}^{E_i+\Delta E/2} dE' \quad
\frac{dn_{\beta}^{\text{IT}}}{dE'} (E')
\end{equation}
where $c$ is the channel index.
Note that the events are binned according to their \emph{reconstructed} energy.
The signal $s_i$ in one bin for a rule now can be composed out of 
several channels by
\begin{equation}
s_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
\end{equation}
where the $\epsilon$'s are weight factors determined by the properties
of the detector.
Similarly the background in one bin $i$ can be specified by
\begin{equation}
b_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
\end{equation}
Quite often it is not sufficient to have constant efficiencies and therefore
it is possible to define efficiencies for each bin and channel $\alpha_i^c$
and thus the event rates in one bin become
\begin{equation}
s_i=\epsilon_{c_1}\alpha_i^{c_1}\cdot n_i^{c_1}\,+
\,\epsilon_{c_2}\alpha_i^{c_2}\cdot n_i^{c_2}\,+\,\ldots\,,
\end{equation}
and analogous for $b_i$.

@@

For the parameterization of an energy calibration error two possibilities
are implemented. The first one being somewhat simpler, whereas the second one
is more accurate but requires a careful choice of parameters. The first 
option is
\begin{equation}
s_i(a,b):=s_i(a)+b\cdot s_i\, E_i/(E_\mathrm{max}-E_\mathrm{min})
\end{equation}
and it is refered to as tilt since it describes a linear distortion, a tilt
of the event rate spectrum. The second option is closer to an actual energy
calibration error, which basically amounts to replacing the reconstructed 
energy $E'$ by $(1+b)\cdot E'$ and we use following approximation
\begin{eqnarray}
s_i(b)&=& (1+b)\cdot(s_{k+1}-s_k)\cdot\delta+s_i\,,\\
\delta&=&b\cdot(i+\Delta E+ 1/2)+i\,,\nonumber\\
k&=&\mod(\delta,1)\,.\nonumber
\end{eqnarray}
Special care is required at the boundaries when $k<1$ or $k+1>N_\mathrm{bins}$,
since there $s_k$ or $s_{k+1}$ may not have been calculated. One possibility
to deal with those cases is to assume that for those values of $k$ $s_k$ is
zero. This is the default. In cases where $s$ at the boundary is however still
sizeable this may lead to intolerable errors and subsequently to a wrong
estimate of the impact of a calibration error. Therefore in those cases it
is advisable to truncate the analysis range by a few bins at the boundary
and therefore ensure in this way that only those $s_i$ are used whose index
$k$ is within the range $1,\ldots, N_\mathrm{bins}-1$.

In order to perform this fine tuning, it is possible to constrain 
the energy range from which the data are taken for the 
analysis\index{Energy window}. This is achieved by setting the energy window
variable. The purpose is mainly to allow a fine tuning of the description
of an energy calibration error. The default setting is that the energy window
coincides with the range defined by the minimal and maximal reconstructed 
energy. For the error dimension using the calibration method (C) it is 
advisable to reduce the energy window by two or three bins at both the lower
and upper end. As a rule of
thumb consider the three standard deviations error on the calibration and
evaluate the corresponding shift in $E'$ at the lower and upper energy limit 
and reduce the analysis range by this amount.

Thus the total event rate $x_i$ in a bin $i$ is given by
\begin{equation}
x_i(a,b,c,d)=s_i(a,b)+b_i(c,d)
\end{equation}
and thus a function of four parameters. The central values for all of
the four parameters are always needed. They are called signal normalization
($a$), signal tilt/calibration ($b$), background  normalization ($c$) and
background tilt/calibration ($d$). The default values are
\begin{equation}
a=1\,,\quad b=0\,,\quad c={\tt NaN}\,,\quad d=0\,,
\end{equation}
thus for the background normalization $c$ a value has to be specified in 
\emph{all} cases.

@@

>>>>

We tell \GLOBES\ that we want to use the built-in neutrino factory
flux originating from stored $\mu^+$'s. This is achieved by setting 
the builtin variable to $1$.  Next the energy of the muons is specified
to be $50\,\mathrm{GeV}$ by the parent energy variable. It is assumed
that there will be $5.33\cdot 10^{20}$ useful muon decays per year
and that this luminosity is available for $8$ years.
\begin{quote}
{\tt /* beam */}\\
{\tt flux(\#mu\_plus)<\\
\tb  @builtin = 1\\
\tb  @parent\_energy = 50.0\\
\tb  @stored\_muons = 5.33e+20\\
\tb  @time = 8.0\\
>}\\
\end{quote}
This specific flux can be later on refered to as {\tt \#mu\_plus}.

\begin{quote}
{\tt \$target\_mass = 50}\\
{\tt \$numofbins = 20}\\
{\tt \$emin = 4.0}\\
{\tt \$emax = 50.0}
\end{quote}
The target mass 
is $50\,\mathrm{kt}$ and a total of $ 4.264\cdot10^{21}$
muons is stored. The data is divided into 20 energy bins ranging from
$4\,\mathrm{GeV}$ to $50\,\mathrm{GeV}$.

Next a file has to be specified which contains the cross section we want to 
use.
\begin{quote}
{\tt /* cross section */}\\
{\tt cross(\#CC)<}\\
{\tt \tb @cross\_file = XCC.dat}\\
{\tt >}
\end{quote}
The command {\tt cross} tells the parser that a cross section environment
begins and has the name {\tt \#CC}. These names are used to refer later on
to this specific environment and each name begins with a leading {\tt \#}.
Whenever we need to specify a cross section from here on we can use {\tt \#CC}
to refer to the cross section as given in the file {\tt XCC.dat}.

@@

The energy resolution function can easily be specified with
\begin{quote}
{\tt /* energy resolution */}\\
{\tt energy(\#MINOS)<}\\
{\tt \tb @type = 1}\\
{\tt \tb @sigma\_e = \{0.15,0.0,0.0\}}\\
{\tt >}
\end{quote}
The {\tt energy} command opens the corresponding energy environment and
the energy resolution function defined here is refered to by the name 
{\tt \#MINOS}. Out of several possibilities it employs algorithm $1$,
the simplest and fastest one. The actual energy resolution is specified
by the energy resolution variable which is a list of three elements. Each 
element is one parameter of a general resolution function as defined in 
equation ~\ref{eq:sigma_e}.

Now we define the one appearance and the corresponding disappearance 
channel of a neutrino factory: 
$\nu_e\rightarrow\nu_\mu$  and $\bar\nu_\mu\rightarrow\bar\nu_\mu$ 
($\mu^+$ stored).
\begin{quote}
{\tt /* channels */}\\
{\tt channel(\#appearance)<}\\
{\tt \tb @channel = \#mu\_plus: +: electron: muon: \#CC: \#MINOS}\\
{\tt >}\\
{\tt channel(\#disappearance)<}\\
{\tt \tb @channel = \#mu\_plus: -: muon: muon: \#CC: \#MINOS}\\
{\tt >}
\end{quote}
The first element is the name of the flux as defined previously, here we use
the flux named {\tt \#mu\_plus}. The second element $\pm$ determines whether 
neutrinos or anti-neutrinos are
taken from the flux table. The third position defines the initial flavour
and the forth position the final flavour, followed by the name of the cross
section and energy resolution function.

The last step is to assemble the channels into a rule.
\begin{quote}
{\tt /* rules */}\\
{\tt rule(\#rule1)<}\\
{\tt \tb @signal = 0.45 @ \#appearance}\\
{\tt \tb @signalerror = 0.001 : 0.0001}\\
{\tt \tb @background = 1.0e-05 @ \#disappearance}\\
{\tt \tb @backgroundcenter = 1 : 0.0}\\
{\tt \tb @backgrounderror = 0.05 : 0.0001}\\
{\tt \tb @errordim = 0}\\
{\tt \tb @energy\_window = 4.0 : 50.0}\\
{\tt >}
\end{quote}
The signal variable says that the signal in our experiment is the 
channel named {\tt \#appearance} with an constant efficiency of $0.45$.
The signal error variable has two components: the first one is the
normalization error of the signal, here it is $0.1\%$. The second component
signifies the energy calibration error of the signal, whose definition 
can be found in section~\ref{sec:energy}. The background variable
specifies the composition of the beam background and this case is
given by the channel named {\tt \#disappearance}, \ie\ this are
the muons whose charge has been mis-identified. They enter the calculation
with an efficiency of $1\cdot 10^{-5}$. The background center variable allows
to rescale the total background contribution which is especially useful if
there are many background components. 
The background error variable is defined in the same way as the signal error
variable, \ie\ we have a $5\%$ background uncertainty.
The error dimension variable selects how
the systematical errors are treated according to table~\ref{tab:error_dim}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Syntax and types of  AEDL}
\label{sec:syntax}

 The first eight characters have to be {\tt \%!GLoBES} 
 this serves to protect from parsing MBytes of chunk
 and cluttering the screen with error messages.

Comments
 are entered like this one -- everything between the 
 delimiters is ignored. 
\begin{quote}
{\tt /* This starts a comment\\
 and here the comment ends */
}
\end{quote}
thus the comments in AEDL work like in C

There are so called  named quantities, they all start with {\tt \$} 
and are predefined. They are checked for being within range, \eg\ {\tt 
\$bins} can be only in between 0 and 150. If you give a {\tt float} where
 an {\tt int} is expected the {\tt float} gets converted to an {\tt int} 
in the usual way in C.  We have named scalar quantities
\begin{quote}
{\tt
\$bins = 10\\
\$baseline = 1200.0
}
\end{quote}
as well as named simple lists
\begin{quote}
{\tt
\$densitytab=\{1.0,2.2343,3.3432\}
}
\end{quote}

Since there are some very special data to parse, some special
provisions have been made so called named environments. This is illustrated 
with the channel definition part
\begin{quote}
{\tt channel(\#ch1)<\\
\tb  $\ldots$\\
>
}
\end{quote}

The first part tells the type of environment, here it is {\tt channel}. 
There are the following types 
\begin{quote}
{\tt flux\\
cross\\
channel\\
energy\\
rule
}
\end{quote}
Besides the type there is a user-defined name beginning with {\tt \#}, in the
above example {\tt \#ch1}, which can be used later on to refer to the 
channel defined in between {\tt <$\ldots$>}. Those names are so called
automatic variables. They start all with {\tt \#} and do not behave as 
usual variables. New names can only be
created within the above construction. Only names which have been created
can be refered to. Using the same name for different types of environments 
does not work.  Each type has its own counter, \eg\ for channel the counter is 
{\tt numofchannels}. The counter keeps track of how many different names 
there are for one type, \ie\ they count the number of channels, 
the number of rules, the number of energy resolution functions asf. 
The value of those automatic variables is the number which can be used 
to refer to them in your C code. You can create a variable for later use by
\begin{quote}
{\tt    channel(\#ch2)<>}
\end{quote}
Now you can refer to {\tt \#ch2} without any trouble and you now could define
a rule using {\tt \#ch2}. Later on of course you still have to specify the
actual channel as described in section~\ref{sec:channel}.

Variables beginning with {\tt @} are named and can be used only within the 
appropriate type of environment, since we have to keep track where 
they belong to. Moreover many of them have a special syntax, like 
{\tt @channel}.

You can have several experiments in one file, just separate the different
 experiments by 
\begin{quote}
{\tt    \#NEXT\#}
\end{quote}
This command resets the counters for channels, rules and energy resolution
functions. All variables have their scope only within a region delimited
by either {\tt \%!GLoBES, \#NEXT\#} or {\tt EOF}. The only exception are 
the counters defined with {\tt cross} and {\tt flux}. They are valid for 
the complete session, since you never must overload the flux and 
cross section information. This ensures  a consistent treatment of various 
experiments in one file - there is no  shadowing of any variables.

Using multiple files to keep things clean and clear is supported by 
 an {\tt include} command. Includes can be nested up 
 to {\tt MAX\_INCLUSION\_DEPTH}, which is set to 10. Error reporting works 
 also for nested includes. The included file is not required to begin 
 with {\tt \%!GLoBES} to facilitate cut and paste
\begin{quote}
{\tt include ./file\_1}
\end{quote}

Thus it is possible and advised to use constructs as the following
\begin{quote}
{\tt    include NuFact.gls\\
        \#NEXT\#\\
        include JHFHK.gls
}
\end{quote}
to define a combined analysis of the experiments defined in the files 
{\tt NuFact.gls} and {\tt JHFHK.gls}. Even if you use the named 
variable {\tt \#CC} in both experiments, but the cross section data used 
is different (\eg\ because of different target nuclei), the correct 
cross section data is used for each of the experiments, since {\tt \#CC} 
in {\tt NuFact.gls} has the value $0$ and
in {\tt JHFHK.gls} it has the value $1$, thus no shadowing occurs.

Furthermore you can store numbers in variables
\begin{quote}
{\tt
Pi = 3.141
}
\end{quote}
And additionally we have some simple calculator functionality like
\begin{quote}
{\tt
Pi+1\\
\verb+Pi^2+\\
sin(Pi/2)\\
}
\end{quote}
The following mathematical functions from {\tt <math.h>} are available 
{\tt sin}, {\tt cos}, {\tt atan}, {\tt ln}, {\tt exp}, {\tt sqrt}. 
Those calculations work everywhere, where
a simple number could stand. However threading over lists does not
 work, \eg\ {\tt sin(\{1,2,3\})} will not work.

One nasty thing is that a line feed character \verb+\n+ is necessary at
 the end of input, alternatively you can put a comment at the end.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Source properties}
\label{sec:source}

\GLOBES\ handles only point sources, thus it is not possible
to study effects from the finite size of the neutrino production
region like in the Sun. Furthermore only one source for each experiment
is possible, this make it impossible\footnote{at least rather tricky} 
to simulate multiple source experiments like KamLAND. All neutrino
sources usable within \GLOBES\ therefore can be specified by the
flux spectra for each flavour and CP sign and the total luminosity
of the source. There are two principal ways to provide a flux:
use a built-in source or a file provided by the user. In both cases
a flux is defined by a named environment.
\begin{quote}
  {\tt flux(\#name)<\\
\tb $\ldots$\\
>}
\end{quote}

In case a built-in source is used the user has to specify which
built-in spectrum is used. Currently two built-in fluxes are available:
$\mu^+$-decay ({\tt @builtin = 1}) and $\mu^-$-decay ({\tt @builtin = 2}).
In any case  the energy of the parent particle has to be specified
and the number of useful decays per year. Thus the sequence for setting
up a neutrino factory is
{\tt flux(\#mu\_plus)<\\
\tb  @builtin = 1\\
\tb  @parent\_energy = 50.0\\
\tb  @stored\_muons = 5.33e+20\\
\tb  @time = 8.0\\
>}\\

Alternatively user defined fluxes may used like this
\begin{quote}
{\tt flux(\#user)<}\\
{\tt \tb @flux\_file = user\_file\_1.dat\\
\tb @time = 2.0\\
\tb @power = 4.0\\
\tb @norm = 1e+8}\\
{\tt >}
\end{quote}

The norm variable allows to set the conversion factor from the fluxes in
the file to the units in \GLOBES. The power variable allows to
specify a target or reactor power. The units used for power are determined
by the units in the flux file and by the value of the norm variable. It is
recommended to use $\mathrm{MW}$ in case a proton based beam is used and
$\mathrm{GW}_\mathrm{thermal}$ for the reactor experiments.

@@

The format of the flux files is seven rows and
501 lines.
\begin{quotation}
$ E\quad
\Phi_{\nu_e}\quad
\Phi_{\nu_\mu}\quad
\Phi_{\nu_\tau}\quad
\Phi_{\bar\nu_e}\quad
\Phi_{\bar\nu_\mu}\quad
\Phi_{\bar\nu_\tau}$
\end{quotation}
where the energy values are evenly spaced. 
For accessing fluxes at 
arbitrary energies linear interpolation is used. Once the energy leaves the
range of values given in the file $0.0$ is returned. The units used for the
fluxes are not specified and right now each flux has its own normalization 
factor as specified by {\tt norm}. The problem is that
in computing the normalization one has to know things like number of target
particles per unit mass asf. which is not always straightforward. The fluxes
will be rescaled by $1/L^2$, thus the normalization must contain a factor
$L_0^2$, when $L_0$ is the distance from the source for which the flux is
given.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Oscillation properties}

@@

The most obvious property of the neutrino detection is the total
active mass of the target as already defined in section~\ref{sec:source} and
its value is defined by 
\begin{quote}
{\tt \$target\_mass = 50.0 }
\end{quote}

@@

A cross section file is read by 
\begin{quote}
{\tt flux(\#name)<}\\
{\tt \tb @cross\_file =user\_file\_1.dat}\\
{\tt >}
\end{quote}  
and can later be refered to by {\tt \#name}.

@@

Since it is rather cumbersome to numerically solve this double integral,
we split the two integrations. We first evaluate the integral over
$\hat{E}$, where the only terms containing $\hat{E}$ are
$k_\beta^{\text{IT}}(E-\hat{E})$,  $ T_\beta(\hat{E})$, and 
$ V_\beta(\hat{E}-E')$, and define:
\begin{eqnarray}
\label{eq:e_res} 
R_\beta^{\text{IT}}(E,E')\,\epsilon_\beta^{\text{IT}}(E')
 \equiv
\int d\hat{E} \quad T_\beta(\hat{E})\,k_\beta^{\text{IT}}(E-\hat{E})
\,V_\beta(\hat{E}-E')\,, 
e\end{eqnarray}

@@

The number of events in one channel and  in one bin\index{Bin} $i$ is given by
\begin{equation}
\label{eq:channel}
n_i^c=\int_{E_i-\Delta E/2}^{E_i+\Delta E/2} dE' \quad
\frac{dn_{\beta}^{\text{IT}}}{dE'} (E')
\end{equation}
where $c$ is the channel index.
Note that the events are binned according to their \emph{reconstructed} energy.
In order to perform the computation \GLOBES\ needs to know the flux,
the CP-sign of the initial state, the initial flavour, the final flavour,
the cross section and the energy resolution function. Employing the notation
of section~\ref{sec:energy} the full formula is

Thus the actual definition of a channel looks like
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
>}
\end{quote}
Furthermore is is possible to define energy dependent efficiencies and
backgrounds (non-beam). These additional components can be either added
before the convolution in with $R(E,E')$ is done, \ie\ the integration with
respect to $E$ in equation~\ref{eq:e_res}. This is called 
{\tt @pre\_smearing\_$\ldots$}. The pre-smearing efficiencies and backgrounds
have to have as many elements as there are sampling points 
(see section~\ref{sec:energy}). In same cases the number of sampling points
is determined automatically by \GLOBES and the number of elements in the 
pre-smearing vectors has to be adjusted in the second run.
Of course, it is possible after all
integrations been performed to have similar factors  and this is called  
{\tt @post\_smearing\_$\ldots$}. All post-smearing factors have to have as 
many elements as there are bins. Efficiencies are multiplicative and their
default value is $1$ whereas backgrounds are additive and thus their default
value is $0$. Thus a more elaborate channel looks like
\begin{quote}
{\tt channel(\#channel\_1)<\\
\tb @channel = \#flux : $+$: muon: muon: \#cross: \#energy\\
\tb @pre\_smearing\_background = \{1,2,3,4,5,6,7,8,9,10\}\\
\tb @post\_smearing\_efficiencies = \{0.1,0.2,0.3,0.4,0.5\}\\
>}
\end{quote}
This experiment obviously uses $10$ sampling points and $5$ bins.

@@

Based on the definition of channels it is possible to assemble the 
signal and background event numbers for each bin.  
The signal $s_i$ in one bin for a rule now can be composed out of 
several channels by
\begin{equation}
s_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
\end{equation}
where the $\epsilon$'s are weight factors determined by the properties
of the detector.
Similarly the background in one bin $i$ can be specified by
\begin{equation}
b_i=\epsilon_{c_1}\cdot n_i^{c_1}\,+\,\epsilon_{c_2}\cdot n_i^{c_2}\,+\,\ldots
\end{equation}
This basic building blocks of a rule are specified by
\begin{quote}
{\tt \tb @signal = 0.5 @ \#channel\_1\\
\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3
}
\end{quote}

Quite often it is not sufficient to have constant efficiencies and therefore
it is possible to define efficiencies for each bin and channel $\alpha_i^c$
and thus the event rates in one bin become
\begin{equation}
s_i=\epsilon_{c_1}\alpha_i^{c_1}\cdot n_i^{c_1}\,+
\,\epsilon_{c_2}\alpha_i^{c_2}\cdot n_i^{c_2}\,+\,\ldots\,,
\end{equation}
and analogous for $b_i$.

@@

For the parameterization of an energy calibration error two possibilities
are implemented. The first one being somewhat simpler, whereas the second one
is more accurate but requires a careful choice of parameters. The first 
option is
\begin{equation}
s_i(a,b):=s_i(a)+b\cdot s_i\, E_i/(E_\mathrm{max}-E_\mathrm{min})
\end{equation}
and it is refered to as tilt since it describes a linear distortion, a tilt
of the event rate spectrum. The second option is closer to an actual energy
calibration error, which basically amounts to replacing the reconstructed 
energy $E'$ by $(1+b)\cdot E'$ and we use following approximation
\begin{eqnarray}
s_i(b)&=& (1+b)\cdot(s_{k+1}-s_k)\cdot\delta+s_i\,,\\
\delta&=&b\cdot(i+t_0+ 1/2)+i\,,\nonumber\\
k&=&\mod(\delta,1)\,,\nonumber\\
t_0&=&E_\mathrm{min}/\Delta E_0\,.\nonumber
\end{eqnarray}
Special care is required at the boundaries when $k<1$ or $k+1>N_\mathrm{bins}$,
since there $s_k$ or $s_{k+1}$ may not have been calculated. One possibility
to deal with those cases is to assume that for those values of $k$ $s_k$ is
zero. This is the default. In cases where $s$ at the boundary is however still
sizeable this may lead to intolerable errors and subsequently to a wrong
estimate of the impact of a calibration error. Therefore in those cases it
is advisable to truncate the analysis range by a few bins at the boundary
and therefore ensure in this way that only those $s_i$ are used whose index
$k$ is within the range $1,\ldots, N_\mathrm{bins}-1$.

In order to perform this fine tuning, it is possible to constrain 
the energy range from which the data are taken for the 
analysis\index{Energy window}. This is achieved by setting the energy window
variable. 
\begin{quote}
{\tt 
\tb @energy\_window = 4.0 : 50.0 
}
\end{quote}
The purpose is mainly to allow a fine tuning of the description
of an energy calibration error. The default setting is that the energy window
coincides with the range defined by the minimal and maximal reconstructed 
energy. For the error dimension using the calibration method (C) it is 
advisable to reduce the energy window by two or three bins at both the lower
and upper end. As a rule of
thumb consider the three standard deviations error on the calibration and
evaluate the corresponding shift in $E'$ at the lower and upper energy limit 
and reduce the analysis range by this amount.

Thus the total event rate $x_i$ in a bin $i$ is given by
\begin{equation}
x_i(a,b,c,d)=s_i(a,b)+b_i(c,d)
\end{equation}
and thus a function of four parameters. The central values for all of
the four parameters are always needed. They are called signal normalization
($a$), signal tilt/calibration ($b$), background  normalization ($c$) and
background tilt/calibration ($d$). The default values are
\begin{equation}
a=1\,,\quad b=0\,,\quad c={\tt NaN}\,,\quad d=0\,,
\end{equation}
thus for the background normalization $c$ a value has to be specified in 
\emph{all} cases. The central values for the normalization and the 
corresponding central value of tilt/calibration are regarded
as a pair. The errors are treated in the same way.
\begin{quote}
{\tt
\tb @signalerror =       0.001  :       0.01\\
\tb @backgroundcenter =  0.1 :       0.0\\
\tb @backgrounderror =   0.001 :       0.01
}
\end{quote}
There is no {\tt @signalcenter} since by default the central value for the
signal normalization is $1$ and the central value for the tilt/calibration 
is $0$.

@@

The error dimensions can be chosen for each rule by
\begin{quote}
{\tt 
\tb @errordim = 1
}
\end{quote}

Next we can assemble the various pieces into one rule
\begin{quote}
{\tt rule(\#rule\_1)<\\
\tb @signal = 0.5 @ \#channel\_1\\
\tb @background = 0.001 @ \#channel\_2 :  0.005 @ \#channel\_3\\
\tb @signalerror =       0.001  :       0.01\\
\tb @backgroundcenter =  0.1 :       0.0\\
\tb @backgrounderror =   0.001 :       0.01\\
\tb @errordim = 1\\
\tb @energy\_window = 4.0 : 50.0\\ 
>}
\end{quote}
